### 2017-06-13 manuscript/manuscript.Rmd
Changes to 'installing and using Git'

Even if you already have Git installed (some computers do) it is a good idea to install the latest version[^git-version], which can be downloaded as a standalone program from <https://git-scm.com/download>. Here, we provide more detailed instructions on installing Git for OS X and Windows, but Linux (and other OS) users can find instructions at the Git website (<https://git-scm.com/book/en/v2/Getting-Started-Installing-Git>). Along with install instructions, we detail how to configure the Git program so that it is ready to be used.

@@

Many OS X computers already have Git installed, especially those operating OS X 10.9 or higher, but it is good practice to install the latest version. The easiest way to update Git to the latest version is to download the installer from <http://git-scm.com/download/mac>, and install it like any other program. Once Git is installed, it can be used from the command line (the OS X command line is available through the Terminal app, which is included by default with the OS X install), or through various GUIs. Although it might be initially intimidating, we encourage using the Terminal app to set the basic configuration of Git.

To operate properly, Git needs to be able to identify its user. You can set these by entering a few basic commands in the Terminal. To verify the current user information, type `git config --global user.name`, and hit return. This should not return anything, unless a previous user of the computer has set it. To ensure that Git knows who you are, type `git config --global user.name "User Name"` (where User Name is your name), and hit return. If you now re-run the first command, Terminal will return the name you entered as `"User Name"`. The second piece of information is your email, which can be entered by `git config --global user.email "email@address.com"` (where email@address.com is your email address). You can verify that the correct email address was saved by typing `git config --global user.email` in the Terminal, and hit return. Once this information is entered, Git will know who you are, and is thus able to track who is doing what within a project, which is especially helpful when you are collaborating with other people.

@@

The first operating principle of Git is that work is organized into independent projects, which Git calls *repositories*. At its core, a repository is a folder on your computer, which is version controlled by Git. Everything that happens inside that folder is tracked by Git, but the user has full control of what is tracked (everything, by default) and when. To establish the operating principles of Git, we now turn to a practical example using a hypothetical project that, we feel, reflects the components of a typical project in experimental psychology.

>>>>

Even if you already have Git installed (some computers do) it is a good idea to install the latest version[^git-version], which can be downloaded as a standalone program from <https://git-scm.com/download>. Here, we provide more detailed instructions on installing Git for OS X and Windows, but Linux (and other OS) users can find instructions at the Git website (<https://git-scm.com/book/en/v2/Getting-Started-Installing-Git>). After installation instructions, we detail how to configure the Git program so that it is ready to be used.

@@

Many OS X computers already have Git installed, especially those operating OS X 10.9 or higher, but it is good practice to install the latest version. The easiest way to update Git to the latest version is to download the installer from <http://git-scm.com/download/mac>, and install it like any other program. Once Git is installed, it can be used from the command line (the OS X command line is available through the Terminal app, which is included by default with the OS X install), or through various GUIs. Although it might at first appear intimidating because of the archaic-looking interface, we encourage using Git from the command line.

To operate properly, Git needs to be able to identify its user. You can set these by entering a few basic commands in the Terminal. The following commands are intended to be entered to Terminal or a similar command line interface. To verify the current user information, type `git config --global user.name`, and hit return. This should not return anything, unless a previous user of the computer has set the global Git user name.

To ensure that Git knows who you are, type `git config --global user.name "User Name"` (where User Name is your name), and hit return. If you now re-run the first command, Terminal will return the name you entered as `"User Name"`. The second piece of information is your email, which can be entered by `git config --global user.email "email@address.com"` (where email@address.com is your email address). You can verify that the correct email address was saved by typing `git config --global user.email` in the Terminal, and hit return. Once this information is entered, Git will know who you are, and is thus able to track who is doing what within a project, which is especially helpful when you are collaborating with other people, or when you are working on multiple computers.

@@

The first operating principle of Git is that your work is organized into independent projects, which Git calls *repositories*. At its core, a repository is a folder on your computer, which is version controlled by Git (you can tell if a folder is "monitored" by git by checking if the folder, or any of its parent folders, contains a hidden `.git` directory). Everything that happens inside a repository is tracked by Git, but the user has full control of what is tracked (everything, by default) and when. Because the user(s) has full control of what and when is tracked, there is a small set of operations the users need to execute every time they wish to log something in Git.

Briefly, when you work in a Git repository, Git stores the state of each file that it monitors, and when any of these files change, Git indicates that they differ from the previously logged state. If the user then is happy with the current changes, she can **add** the changed files to Git's "staging area". If the user then is certain that the files added to the staging area are in good shape, they can **commit** the changes. These two operations are the backbone of using Git to store the state of the project whenever meaningful changes are made.

Most importantly, each state in Git's log contains the full state of the files at that point in time. Users can always go back to an earlier version by "checking out" a previous state of Git's log. That's why these programs are called Version Control Systems.

To establish the operating principles of Git in practice, we now turn to a practical example using a hypothetical project that, we feel, reflects the components of a typical project in experimental psychology.
